---
title: "Class 9 Week 5 Mon Apr 23 Defensive"
author: "david"
date: "April 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**DEFENSIVE PROGRAMMING & DEBUGGING**

2 ways: style guide (documentation) and failing fast (meaningful error message)

##Style
- give variables, functions, and files *short* *intuitive* names that describe what they do
- NO spaces NO caps NO periods [separate by underscores]
- do not create names that over-write objects you did not create [i.e. T, F, mean]

- do NOT add explanations of results of a graph or table as a #comment, add as text to document outside of chunk

*nitpicking*
-When packages have identical names for different functions, use double colons to call forth the functions
dplyr::filter() vs. MASS::filter()
-use spaces between objects, etc
-indent after functions/pipes
-no run-on lines [dont make them scroll horizontally on .md files; use return key]
-comments on separate lines:
   - code folding operations...?

*code reformatting: (dropdown menu)* 
these auto-solve the style nitpick issues the code:

Code > Reformat Code (Shift + Cmd/Ctrl + A) 
  splits into lines, adds spaces

Code > Reindent Lines (Cmd/Ctrl + I)
  auto-indents inside pipes/functions


#Exercise
```{r}
library(tidyverse)
library(modelr)
library(broom)
library(gam)
College <- as_tibble(ISLR::College)%>%mutate(Outstate =Outstate/1000,Room.Board=Room.Board/ 1000)# rescale Outstate in thousands of dollars
crossv_kfold(College,k=10)%>%mutate(linear=map(train,~glm(Outstate~PhD, data=.)),log= map(train,~glm(Outstate ~log(PhD), data=.)),spline=map(train,~ glm(Outstate ~bs(PhD, df=5),   data=.)))%>%gather(type,model,linear:spline)%>%mutate(mse=map2_dbl(model,test,mse))%>%group_by(type)%>%summarize(mse = mean(mse))# k-fold cv of three model types
college_phd_spline<-gam(Outstate~bs(PhD,df=5),data=College)# spline has the best model fit
college_phd_terms<-preplot(college_phd_spline,se=TRUE,rug=FALSE)# get first difference for age
#age plot
data_frame(x=college_phd_terms$`bs(PhD, df = 5)`$x,
y=college_phd_terms$`bs(PhD, df = 5)`$y,
se.fit = college_phd_terms$`bs(PhD, df = 5)`$se.y)%>%
mutate(y_low = y - 1.96 * se.fit,y_high = y+1.96 * se.fit) %>%ggplot(aes(x, y))+geom_line()+
geom_line(aes(y = y_low), linetype = 2)+
geom_line(aes(y = y_high), linetype = 2)+
labs(title = "Cubic spline of out-of-state tuition",subtitle = "Knots = 2",x = "Percent of faculty with PhDs",y=expression(f[1](PhD)))
```


#solution
```{r exercise_style}

library(tidyverse)
library(modelr)
library(broom)
library(gam)

College <- as_tibble(ISLR::College) %>% 
# rescale Outstate in thousands of dollars
  mutate(Outstate = Outstate / 1000, 
         Room.Board = Room.Board / 1000)

# k-fold cv of three model
crossv_kfold(College, k = 10) %>% 
  mutate(linear = map(train,  ~ glm(Outstate ~ PhD, data = .)),
         log = map(train,  ~ glm(Outstate ~ log(PhD), data = .)),
         spline = map(train,  ~ glm(Outstate ~ bs(PhD, df = 5),   data = .))) %>% 
  gather(type, model, linear:spline) %>% 
  mutate(mse = map2_dbl(model, test, mse)) %>%
  group_by(type) %>% 
  summarize(mse = mean(mse)) 

# spline has the best model fit
college_phd_spline <- gam(Outstate ~ bs(PhD, df = 5), data = College)

# get first difference for age
college_phd_terms <- preplot(college_phd_spline, se = TRUE, rug = FALSE)

#age plot
data_frame(x = college_phd_terms$`bs(PhD, df = 5)`$x,
           y = college_phd_terms$`bs(PhD, df = 5)`$y,
           se.fit = college_phd_terms$`bs(PhD, df = 5)`$se.y) %>%
  mutate(y_low = y - 1.96 * se.fit, 
         y_high = y + 1.96 * se.fit) %>% 
  ggplot(aes(x, y)) +
    geom_line() +
    geom_line(aes(y = y_low), linetype = 2) +
    geom_line(aes(y = y_high), linetype = 2) +
    labs(title = "Cubic spline of out-of-state tuition",
         subtitle = "Knots = 2",
         x = "Percent of faculty with PhDs",
         y = expression(f[1](PhD)))


```


#**DEBUGGING**

conditions= communicating that something unexpected occurred [3 types]

- (1) errors (fatal) 
      code stops bc R doesnt know what to do
          e.g if(x < 0) {stop('x not < 0')}... 
          what if we want the function to continue, but spit out NA? use '(2) warning'
          or use try({your.function}) to ignore errors and do it anyway????
          
- (2) warnings
          function continues to run, produces the output, but tells you what happened
      

- (3) messages
        nothing went wrong, but tells you what happened in the code
        how to suppress message? use suppressMessages(yourfunction())
        *can have function(x) message("message thooo")
        *use the message() function to generate message, not print()
        *if you dont know what generated the message, wrap ggplot function:
        suppressMessages({
             ggplot() +
             geom_point() +
             geom_smooth()
          })
          
#function with condition excercise (1. make throw an error when negative sqrt; 2. make throw warning + output NA)
```{r}
sqrt_sum <- function(x, y){
  if(sum(x, y) < 0) stop("The sum of these numbers is negative.")
  sqrt(x + y)
}

sqrt_sum(3, -6)
```

#solution 1.
```{r}
sqrt_sum <- function(x, y){
  if(sum(x,y) < 0) {stop("You're tearing me apart, Lisa!")}
  else {sqrt(x + y)}
}

sqrt_sum(3, -6)

```

#solution2.
```{r}
sqrt_sum <- function(x, y){
  if(sum(x,y) < 0) {message("You're tearing me apart, Lisa!") & return(NA)}
  else {sqrt(x + y)}
}

sqrt_sum(3, -6)
          
```

- problemn: how to figure out which function in chunk is causing the error
traceback()... read output from bottom to top (bottom=function you called)(top=cause of error message)

*condition handling (stop errors): 2 options*

try()
   a function you can use to ignore errors. wrap the function in try(): try(thefunction)
   if try() does not succeed, just goes on to next line of code, doesn't stop & throw error (silently ignores)
   
safely() 
    its an adverb... modifies other function verbs to behave safely/consistantly
    returns a *list* with 2 vectors for each element
    - result (if operation was successful, there will be a result
              if operation fails, there will be a NULL value
    - error (if operation fails, there will be an error message 
              if succeeds, value=NULL)
    great bc very predictable output, gives you a record of what happened for each element/iteration
    *transpose() it like safely(), but more complex ... can extract the values that returned errors or the 
    values that returned results (result[is_ok] or with a !)
    flatten_dbl() --> flatten from list into vector






