---
title: "Class 6 Apr 11 Week 3 Wed"
author: "david"
date: "April 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Relational Data*
not just one table, combination, relations between multiple data frames encode information
Relations are between two tables only, can relate a->b then ab->c but not a->b<-c

relational verbs:
mutating joins: create new columns in table
filter values in table


#join functions

e,g, superheroes and publishers data frames: 
why 2? 
observations are on completely different type of thing, dont belong in same table (charachters in 1, publishers in 2)
how to join? mutating join: both tables have "publisher" column
```{r}
inner_join(x = superheroes, y = publishers)
# x= , y= put names of two tables
```
hellboy is lost bc its publisher isnt in y

left_join(x=x, y=y) keeps all rows from x dataframes, and keeps all columns from x and y dtaframes; inner_join drops all rows that dont have matches

right_join returns all rows from y, all columns from x and y

full_join keeps all rows and columns from x AND y

Merge function is base R function: it can do all of these kinds of joins,
can write inside functions
in dplyr, put which join you want to do as an argument inside functions

R looks for key columns itself: assumes columns with same name in both datafraames is key column and contains information
another arguemnt "by=" can tell R which columns in dataframes 1 and 2 are key

#filter join functions 

semi_join(x=1stdataframe,y=2nddataframe) only keeps columns that were in 1st data frame..
uses 2nd dataframe to filter 1st dataframe, doesnt add any add'l values from 2nd

anti_join(x=,y=) keeps rows in x that do NOT have a corresponding row in y

#exercises 

library(tidyverse)
library(nycflights13)
theme_set(theme_minimal())


age and departure: join planes with flights
one row for each flight: keep all rows in flights

```{r}
flights<-mutate(flights, yearflight = year)
tailflight<-select(flights, tailnum, yearflight, dep_delay)
age_delay<-left_join(x=tailflight, y=planes)
age_delay2 <- mutate(age_delay, year = yearflight-year)
ggplot(age_delay2, aes(year, dep_delay)) +
  geom_smooth()
```

```{r}
plane_age <- planes %>%
  mutate(age = 2013-year) %>%
  select(age, tailnum)

left_join(x=plane_age, y=flights) %>%
  ggplot(aes(age, dep_delay)) +
    geom_smooth()
```

#Transforming and Visualizing factors
factors used to store categorical values, eg strings of text, store associated labels
can conver cha. var to factor var when there is FINITE amount of values a var could take on (fixed num of categories with specific assoc labels)

```{r}
library(tidyverse)
library(knitr)
library(rcfss)
theme_set(theme_minimal())

# load the data
data("gun_deaths")
gun_deaths
```

```{r}
spergajk<-gun_deaths%>%
  mutate(month = factor(month, levels = c(1:12), labels = month.abb))

spergajk %>%
  ggplot(aes(month)) +
  geom_bar() +
  labs(x="Month", y="# of deaths", title="U.S. Gun Deaths (2012-2014), by Month")

spergajk %>%
  count(month) %>%
  ggplot(aes(fct_reorder(month, n), n)) +
  geom_col() +
  labs(x="Month", y="# of deaths", title="U.S. Gun Deaths (2012-2014), by Month")

spergajk %>%
  count(month) %>%
  ggplot(aes(fct_reorder(month, n), n)) +
  geom_col() +
  labs(title = "Gun Deaths in the United States (2012-2014)",
       x = "Month",
       y = "Number of gun deaths")
 
```
