---
title: "Class 8 Week 4 Wednes April 18"
author: "david"
date: "April 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#week 4 wednes
*Week 6 Monday at bottom*

vectors and iteration tho
dataframe is not the basic unit of data: its VeCtOrS!

2 types
Atomic Vectors - homogenous, 1 type of data only
List Vectors (lists) - heterogenous, flexible with many types

NA = Null (null means missing vector as opposed to na meaning missing value)44

logical vector: either true or false or missing value
BUUUH<-c(TRUE, FALSE, TRUE, BA)

Numeric vectors: most numbers stored as double values; with decimal points... they contain numbers
c(1.3,4.2,5)
(integers can only do whole numbers)

Character vectors [its in quotes!]
c("goodnight","hello","buuu")

A scalar is a vector of length one!
if you add a scalar to a vector, each value of the vector increases by the scalar
100+ c(1,2) = c(101,102)

problem: adding vectors of different length: will repeat the smaller vector again and again until lengths are same (eggu eggu eggu eggu)


*** filter function works for dataframes, NOT vectors
if you want to *SUBSET* vector
- use positive integers (to indicate the position of the value in the prior vector that you select)
      *extract from x vector the elements at the following positions
- use negative integers
      *extract from x vector all EXCEPT those at the following positions
x=("a","b","c")
x[c(3,2)]
will output: "c" "b"
cannot mix positive and negative integers

- can filter by missing/non-missing values
*keep only missing values
x[is.na(x)]
*keep only non-missing values
x[!is.na(x)]

- can filter by values of elements
*keep only values > 5
x[x >5]

can use modular division
x[x %% 2 ==0] #keep all elements in X that are divisible evenly by 2 (no remainder) 
%%=modular division
^this does not filter out missing values

```{r}
(x <- seq(from = 1, to = 10))
x[c(1:4,7)]
```

```{r}
(x <- seq(from = 1, to = 10))
x[-9] #or x[c(-9)]
```

```{r}
(x <- seq(from = 1, to = 10))
x>5
x[x >5]
```

```{r}
(x <- seq(from = 1, to = 10))
x[x %% 3 == 0]
```

LISTS
a dataframe is a special type of list
- in list, each row is a vector
- all list columns must have same length
- subset of list will alwats be another list
a[[1]] --> list subset to vector (strip away the list hierarchy, extract the element, a vector)
a[1] --> list subset to list
- eg, element of list = row = vector


```{r}
z <- list(a = 1:3, b = "a string", c = pi, d = list(-1, -5))
str(z)
z[["a"]]
z[pi]
z[c(1,3)]

```

ITERATIVE OPERATIONS
for loop

3 elements
1. output (where ami gonna stooore)
2. sequence (what am i looping over)
3. operation (what the function actually does)

1.
output <- vector(mode="double", length=ncol(df))

2.
i in seq_along(df)
seq_along(df)
## [1] 1 2 3 4

first time through loop, i will equal 1, 2nd time, i=2... etc

3.
```{r}
output[[i]] <- median(df[[i]]) # we are calculating median of the i'th column (1st, 2nd etc)
#1st, 2nd, 3rd, 4th column median values are calculated and stored in succession
```

```{r}
#complete for loop:
output <- vector(mode = "double", length = ncol(df))
for (i in seq_along(df)) {
  output[[i]] <- median(df[[i]])
}
```
#preallocation (assign output) increases speed of computation

```{r}
#calculate mean for every column
library(tidyverse)
library(rcfss)
library(knitr)
as_tibble(mtcars)
output <- vector(mode="double", length=ncol(mtcars))
for (i in seq_along(mtcars)) {
  output[[i]] <- mean(mtcars[[i]])
}
output


```


MAP FUNCTIONS
part of the library(purr) package of tidyverse.. verbs>>>nouns
typles of map() functions

map() makes a list
map_lgl() makes a logical vector
map_int() makes an integer vector
map_dbl() makes a double vector
map_chr() charachter vector

map_dbl(dataframe,mathematical operation you want to do on each column, na.rm= TRUE)

eg
map_dbl(mtcars, mean, na.rm=TRUE)

can be piped tho
 
SCOPED VERBS
3rd kind of iterative operation:
- input is always dataframe, output is always dataframe

mtcars %>%
  summarize() or mutate() or filter()

which vars to operate on?

_if
pick vars to operate on based on another function (eg if a charachter variable)
mtcars %>%
summarize_if(is.numeric, mean, is.na=TRUE)


_at
pick specific variables like select function that you name
summariz_at(mtcars,vars(mpg),funs(min,max))

funs=which operations you do
vars=which vars get operated on

_all
all vars

summarize_all(mtcars, funs(min, max))

*_*_*_*_*_*
Helpful for map() functions:

**can use a nest() function to create a list column in the dataset, where each value is a dataframe
**use a broom function

y_country <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()

by_country
## # A tibble: 142 x 3
##        country continent              data
##         <fctr>    <fctr>            <list>
##  1 Afghanistan      Asia <tibble [12 x 4]>
##  2     Albania    Europe <tibble [12 x 4]>
##  3     Algeria    Africa <tibble [12 x 4]>
##  4      Angola    Africa <tibble [12 x 4]>
##  5   Argentina  Americas <tibble [12 x 4]>
##  6   Australia   Oceania <tibble [12 x 4]>
##  7     Austria    Europe <tibble [12 x 4]>
##  8     Bahrain      Asia <tibble [12 x 4]>
##  9  Bangladesh      Asia <tibble [12 x 4]>
## 10     Belgium    Europe <tibble [12 x 4]>
## # ... with 132 more rows